SIR Model with Testing and Isolation Mechanisms

### 2020-11-02 codes/ali_test_sir.Rmd
updated testing strategies analysis

1- What combination of testing intensity, $\rho$, and isolation, $\eta_w$ and $\eta_t$, will result in $R_0 <1$?

@@

**Notes:** 
- The complicated part is the flow rate from $Z_u$ to $Z_n$, $Z_p$. The (total, not per capita) flow from $Z_u$ to $Z_n$ is $\rho (1-P_z) F(W_z) Z_u$, from $Z_u$ to $Z_p$ is $\rho P_Z F(W_Z) Z_u$. Where, $P_z$ is the probability of positive test. Thus, $P_i$ represents the test accuracy. We assume $P_s=0$.
- On testing and scaling; at the simplest $F_z=W_z$.  
- JD: If this $\tau$ is fast (e.g., 1d), it should have minimal effects on the system when we're in the usual domain. Since $\tau \gg \rho$, this generally collapses to the original form. When $W$ is super-small, however, it collapses instead to $\tau$.

@@

```{r model inputs, message=FALSE}
params <- c(N0=1000000, beta=0.8,gamma=1/3, omega=0.25, rho=0.01, 
             W_s=0.01, W_i=1, W_r=0.01, #weights
             P_s=0, P_i=1, P_r=0.5, #prob of waiting for being tested positive
            eta_w=0.02, eta_t=0.01, #isolation parameter 
            s=2, #eta_w/eta_t
            tau=1
            )
class(params) <- "params_pansim"
state_init <- c(S_u=params[["N0"]], S_n=0,
                I_u=1,I_n=0,I_p=0,I_t=0,
                R_u=0,R_n=0,R_p=0,R_t=0,
                N=0,P=0)
d <- seq(0,200,by=0.1)
# Conditional stop for desolver
rootfun <- function (time, state,params) { 
  unpack(as.list(c(state,params)))
  return(I_u - 1) }
```

@@

```{r run2, message=FALSE}
d_long <- seq(0,300,by=0.1)
## capturing badly behaved integrator warning messages
cc <- capture.output(
    out_long <- as.data.frame(
        ode(
    func=sir.model,
    y=state_init,
    times= d_long, 
    parms=update(params,beta=2)
    # rootfun = rootfun,
    # method="lsodar"
  )
  ))
out_long2 <- out_long %>%
    pivot_longer(c(S_u,S_n,I_u,I_n,I_p,I_t, R_u,R_n,R_p,R_t),
                 names_to = "compartment", values_to = "value")

@@

R0<-function(params){
    unpack(as.list(params))
    Su_dfe<- Su_dfe(params)
    Sn_dfe<- N0-Su_dfe
    # print(c(Su_dfe,Sn_dfe,Su_dfe+Sn_dfe))
    F_i <- F_i(params) #at DFE

    A <-gamma*(omega+gamma+eta_w*F_i) + omega*eta_t*P_i*F_i
    B <-(omega+eta_w*(F_i+gamma))*gamma+(eta_w*gamma+eta_t*omega)*(omega*P_i*F_i)/(omega+gamma)

    C <- beta/(N0*gamma*(gamma*(omega+gamma)+F_i*(gamma+omega*P_i)))
    return((A*Su_dfe+B*Sn_dfe)*C)
}

@@

### Contour plot of $R_0=1$ and interpretations approache 1
```{r contour_R0_1, message=FALSE,warning=FALSE }
n_out <- 5 ## facets (rows/cols)
n_in <- 41 ## grid N within facets
 # specify the ranges
 beta  <- 0.4  ## set beta high enough to allow R0>1 in worst case
 eta_w <- seq(0,1, length.out=n_out)
 eta_t <- seq(0.001,0.5, length.out=n_out) #0.001 to 0.5 
 rho     <- seq(0,0.01, length.out=n_in)
 omega <- seq(0.001,1 , length.out=n_in) #note omega must be > rho, was 0.1

@@

```{r contour_R0_2, message=FALSE,warning=FALSE }
 # specify the ranges
 beta  <- 0.4  
 # eta_w <- seq(0,1, length.out=5) 
 eta_t <- seq(0.001,0.5, length.out=n_out) #0.001 
 rho     <- seq(0,0.01, length.out=n_in)
 omega <- seq(0.001,1 , length.out=n_in) #note omega must be > 0, was 0.1
 s <- seq(1,2,length.out=n_out)

@@

df22<- data.frame(df11,R0=apply(df11,1,function(params_in)R02(params=params_in)),
                      Fi= apply(df11,1,function(params_in)F_i(params=params_in)),
                       Su=apply(df11,1,function(params_in)Su_dfe(params = params_in)))
min(df22$Su)
which(df22[]<0)          
```

@@

>>>>

- An exploration of simple mathematical/computational models of COVID-19 to gain insight into the effects of different levels of testing and isolation on $R_t$/growth rates. The goal is insight.

@@

**Notes:**

- The complicated part is the flow rate from $Z_u$ to $Z_n$, $Z_p$. The (total, not per capita) flow from $Z_u$ to $Z_n$ is $\rho (1-P_z) F(W_z) Z_u$, from $Z_u$ to $Z_p$ is $\rho P_Z F(W_Z) Z_u$. Where, $P_z$ is the probability of positive test. Thus, $P_i$ represents the test accuracy. We assume $P_s=0$.

- On testing and scaling; at the simplest $F_z=W_z$.

- JD: If this $\tau$ is fast (e.g., 1d), it should have minimal effects on the system when we're in the usual domain. Since $\tau \gg \rho$, this generally collapses to the original form. When $W$ is super-small, however, it collapses instead to $\tau$.

### On Testing

- Initial approach and singularity issue (not sure if I call it structural or numerical singularity?); 
$\sigma = \frac{\rho N_0}{W}$ was used initially. The issue was that the population in $S$ compartments appeared to blow up when the DFE is achieved ( in the simulation when time, $t>200$). This is  once the only untested people are susceptibles, the FoI will become $\Lambda=0$, testing rate $F_s=t N_0/S_u$. Thus, eq(1) of the model will be $d S_u/dt = - t N_0 + \omega S_n$ which is no longer dependent on $S_u$ and a linear rate of leaving the $S_u$ compartment.

@@

```{r model inputs, message=FALSE}
params <- c(N0=1000000, beta=0.8,gamma=1/3, omega=0.25, rho=0.01, 
             W_s=1, W_i=1, W_r=1, #weights
             P_s=0, P_i=1, P_r=0.5, #prob of waiting for being tested positive
            eta_w=0.02, eta_t=0.01, #isolation parameter 
            s=2, #eta_w/eta_t
            tau=1
            )
class(params) <- "params_pansim"
state_init <- c(S_u=params[["N0"]], S_n=0,
                I_u=1,I_n=0,I_p=0,I_t=0,
                R_u=0,R_n=0,R_p=0,R_t=0,
                N=0,P=0)
d <- seq(0,200,by=0.1)
# Conditional stop for desolver
rootfun <- function (time, state,params) { 
  unpack(as.list(c(state,params)))
  return(I_u - 1) }
```

@@

```{r run2, message=FALSE}
d_long <- seq(0,400,by=0.1)
## capturing badly behaved integrator warning messages
cc <- capture.output(
    out_long <- as.data.frame(
        ode(
    func=sir.model,
    y=state_init,
    times= d_long, 
    parms=update(params,beta=2)
    # rootfun = rootfun,
    # method="lsodar"
  )
  ))
out_long2 <- out_long %>%
    pivot_longer(c(S_u,S_n,I_u,I_n,I_p,I_t, R_u,R_n,R_p,R_t),
                 names_to = "compartment", values_to = "value")

@@

R0<-function(params){
    unpack(as.list(params))
    Su_dfe<- Su_dfe(params)
    Sn_dfe<- N0-Su_dfe
    # print(c(Su_dfe,Sn_dfe,Su_dfe+Sn_dfe))
    Fi <- F_i(params) #at DFE

    A <-gamma*(omega+gamma+eta_w*Fi) + omega*eta_t*P_i*Fi
    B <-(omega+eta_w*(Fi+gamma))*gamma+(eta_w*gamma+eta_t*omega)*(omega*P_i*Fi)/(omega+gamma)

    C <- beta/(N0*gamma*(gamma*(omega+gamma)+Fi*(gamma+omega*P_i)))
    return((A*Su_dfe+B*Sn_dfe)*C)
}

@@

### Contour plot of $R_0=1$ and interpretations approache 1
```{r contour_R0_1, message=FALSE,warning=FALSE }
n_out <- 5 ## facets (rows/cols)
n_in <- 41 ## grid N within facets
 # specify the ranges
 beta  <- 0.5  ## set beta high enough to allow R0>1 in worst case
 eta_w <- seq(0,1, length.out=n_out)
 eta_t <- seq(0.001,0.5, length.out=n_out) #0.001 to 0.5 
 rho     <- seq(0,0.01, length.out=n_in)
 omega <- seq(0.001,1 , length.out=n_in) #note omega must be > rho, was 0.1

@@

```{r contour_R0_2, message=FALSE,warning=FALSE }
 # specify the ranges
 beta  <- 0.5
 # eta_w <- seq(0,1, length.out=5) 
 eta_t <- seq(0.0001,0.5, length.out=n_out) #0.001 
 rho     <- seq(0,0.01, length.out=n_in)
 omega <- seq(0.001,1 , length.out=n_in) #note omega must be > 0, was 0.1
 s <- seq(1,2,length.out=n_out)

@@

df22<- data.frame(df11,R0=apply(df11,1,function(params_in)R02(params=params_in)),
                      Fi= apply(df11,1,function(params_in)F_i(params=params_in)),
                       Su=apply(df11,1,function(params_in)Su_dfe(params = params_in)))
# min(df22$Su)
# which(df22[]<0)          
```

@@

4. Taylor expansion gives $R_0|_{(\rho=0,W_s=1)} = \beta/\gamma$.

5. Taylor expansion gives $R_0|_{(\rho=0,W_s=0)} = \beta/\gamma \frac{\gamma \tau(\omega+\gamma)W_s+\tau\rho(\gamma\eta_w+\omega\eta_tP_i)W_i+ \gamma \rho(\gamma+\omega)}{\gamma \tau(\omega+\gamma)W_s+\tau\rho(\gamma+\omega P_i)W_i+ \gamma \rho(\gamma+\omega)}$.

### 2020-11-02 note/SIR_notes.Rnw
testing function is updated, code is updated not the text
updated testing strategies analysis

% %%%%%%%
\subsection{Other maybe-related works}
\citep{arino2020simple} developed a SLIAR compartmental model to study the spread of an epidemic, specifically COVID-19, in a population. The model incorporates an Erlang distribution of times of sojourn inincubating, symptomatically and asymptomatically infectious compartments. Basic reproduction number is derived. Also, sensitivity analysis with respect to the underlying parameters for the following two outputs was carried out; (i) the number of observable cases during the course of the epidemic and at the peak, and (ii) the timing of the peak of the outbreak. Sensitivity analysis is performed using the R package multisensi.
% %%%%%%%
\bibliography{../SIRlibrary}
\end{document}

>>>>

\citep{tuite2020mathematical} used an age-structured compartmental model of COVID-19 transmission in the population of Ontario, Canada. We compared a base case with limited testing, isolation and quarantine to different scenarios.

% %%%%%%%
\subsection{Other maybe-related works}
\citep{arino2020simple} developed a SLIAR compartmental model to study the spread of an epidemic, specifically COVID-19, in a population. The model incorporates an Erlang distribution of times of sojourn inincubating, symptomatically and asymptomatically infectious compartments. Basic reproduction number is derived. Also, sensitivity analysis with respect to the underlying parameters for the following two outputs was carried out; (i) the number of observable cases during the course of the epidemic and at the peak, and (ii) the timing of the peak of the outbreak. Sensitivity analysis is performed using the R package multisensi.

\citep{ruszkiewicz2020diagnosis} novel with-in-a-minute breath testing with 80\% accuracy. 
% %%%%%%%
\bibliography{../SIRlibrary}
\end{document}
